|

### 2020-06-03 paper.Rmd
More on regularisation, sigma rescaling.

- linear projections and tours [@As85,@BCAH05]: we want to discover features, by looking at the data from all sides
- most projections are approximately gaussian, as dimensionality increases, observations pile in the center [@diaconis1984]
- related: crowding problem as addressed in t-SNE [@tsne]
- connection with high-dimension low-sample data? observations will be away from the center, on the vertices of a simples [@doi:10.1111/j.1467-9868.2005.00510.x], and we can find directions where data is piling in a single point [@10.2307/27639976,@10.1093/biomet/asp084]
- connection with regularisation? (optimisation vs visualisation)

@@

## Smudging: rescaling the radius

@@

@@

# plot dependence for selected values
ggplot(data = data.frame(x = c(0,1)), mapping = aes(x = x)) +
    stat_function(
        fun = trans_sig(1),
        mapping = aes(color = "ca")) +
  stat_function(
        fun = trans_sig(0.5),
        mapping = aes(color = "cb")) +
  stat_function(
        fun = trans_sig(0.3),
        mapping = aes(color = "cc")) +
    stat_function(
        fun = trans_sig(0.2),
        mapping = aes(color = "cd")) +
      stat_function(
        fun = trans_sig(0.1),
        mapping = aes(color = "ce")) +
  scale_color_manual(name = expression(sigma),
                     values = RColorBrewer::brewer.pal(5, "Dark2"),
                     labels = c("1 (baseline)","0.5", "0.4", "0.2", "0.1")) +
  xlab("r") + ylab("r'") + theme_bw()

@@

Things to note here: rescaling is nearly linear near the center of the distribution (that we want to focus on), limited distortion here. "Curlig" effect near the maximum radius

@@

What if we assume a multivariate normal instead? Maybe this is actually not that different?

>>>>

- linear projections and tours [@As85,@BCAH05]: we want to discover features, by looking at the data from all sides
- most projections are approximately gaussian, as dimensionality increases, observations pile in the center [@diaconis1984]
- related: crowding problem as addressed in t-SNE [@tsne]
- connection with high-dimension low-sample data? observations will be away from the center, on the vertices of a simples [@doi:10.1111/j.1467-9868.2005.00510.x], and we can find directions where data is piling in a single point [@10.2307/27639976,@10.1093/biomet/asp084]
- for modelling, we often address high-dimensional problems with regularisation methods (ridge regression, Lasso). These are optimisation techniques that shrink the parameter estimates (ridge regression gives global shrinkage, Lasso can be thought of as variable selection and will shrink some of the estimates to zero). Our approach is complementary, we still consider all variables and possible combinations, but move the focus away from the extreme values, instead focusing in towards the center, where observations tend to pile up.

We have to think about the low-d projections differently than the full distribution. As $p$ increases the full volume of the parameter space grows exponentially, and we get a low density of points, and in general the center of the distribution will be empty [@doi:10.1111/j.1467-9868.2005.00510.x]. Similarly, sections of the data will typically capture only a very small fraction of the points. When making projections, we are effectively marginalising the distribution over all orthogonal directions, and this averaging often leads to the piling of observations in the center.

XXX maybe in real applications we need to do both? Use some form of dimension reduction that can screen some of the noise, followed by zooming in on the center?

@@

## Rescaling the radius

@@

XXX as $\sigma$ approaches $r/2$ we see that $r'_{max} < r_{max}$, and larger values of $\sigma$ do not make sense. Using $\sigma=r$ gives linear scaling that shrinks the overall scale...

@@

# plot dependence for selected values
ggplot(data = data.frame(x = c(0,1)), mapping = aes(x = x)) +
    stat_function(
        fun = trans_sig(0.4),
        mapping = aes(color = "ca")) +
  stat_function(
        fun = trans_sig(0.3),
        mapping = aes(color = "cb")) +
  stat_function(
        fun = trans_sig(0.2),
        mapping = aes(color = "cc")) +
    stat_function(
        fun = trans_sig(0.1),
        mapping = aes(color = "cd")) +
      stat_function(
        fun = trans_sig(0.05),
        mapping = aes(color = "ce")) +
  scale_color_manual(name = expression(sigma),
                     values = RColorBrewer::brewer.pal(5, "Dark2"),
                     labels = c("0.4", "0.3", "0.2", "0.1", "0.05")) +
  xlab("r") + ylab("r'") + theme_bw()

```{r circles-normal, fig.cap="Rescaled equidistant concentric circles."}
library(ggforce)

lbl <- c("sig=0.3", "sig=0.25", "sig=0.2", "sig=0.15")
names(lbl) <- c("s03", "s025", "s02", "s015")

tibble(x = 0, y = 0, r = seq(0.1, 1, length.out = 10)) %>%
  mutate(s03 = trans_sig(0.3)(r)) %>%
  mutate(s025 = trans_sig(0.25)(r)) %>%
  mutate(s02 = trans_sig(0.2)(r)) %>%
  mutate(s015 = trans_sig(0.15)(r)) %>%
  mutate(r = as.factor(r)) %>%
  pivot_longer(cols = starts_with(("s"))) %>%
  mutate(name = factor(name, levels = c("s03", "s025", "s02", "s015"))) %>%
  ggplot() +
  geom_circle(aes(x0=x, y0=y, r=value, color=r)) +
  coord_fixed() +
  theme_bw() +
  facet_wrap(~name, labeller = labeller(name = lbl))

@@

Things to note here: rescaling is nearly linear near the center of the distribution (that we want to focus on), limited distortion here. "Curling" effect near the maximum radius

@@
